---
title: "Assignment-3"
author: '******'
date: "May 29, 2018"
output:
  pdf_document: default
  html_document: default
---
#Introduction#

Objective of the study was to create a dataset which could be used to predict the relationship between Number of accidents and the total vehicle ownership in Australian states for the year 2016.

For the analysis two datasets were considered

1.Data describing the number of crashes and variables which explains the scenarion in which crashes occured - https://bitre.gov.au/statistics/safety/fatal_road_crash_database.aspx

2.Number of vehicle ownership in Australian States
http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/9208.012%20months%20ended%2030%20June%202016?OpenDocument

#Importing Data in R#




#Checking the Dimensions $ class of objects#

```{r}
install.packages("readr")
install.packages("gdata")
library(readr)
library(gdata)
Road_Fatalities <- read.csv("Road_Fatalities.csv", stringsAsFactors = FALSE)
str(Road_Fatalities)
Total_Vehicles_and_Km <- read.csv("Total_Vehicles_and_Km.csv")
str(Total_Vehicles_and_Km)
Total_Vehicles_and_Km
```

All the unavaialbale data is coded as "-9", which will be handled in missing data handling in the upcoming sections.


#Assigning Labels#

The "State" variable is character type by default and this to be converted to factor variable
```{r}
Total_Vehicles_and_Km$Country<-factor(Total_Vehicles_and_Km$Country,levels=c("New South Wales", "Victoria","Queensland","South Australia","Western Australia","Tasmania","Northern Territory","Australian Capital Territory"),
                                      labels = c("NSW", "VIC","QLD","SA","WA","TAS","NT","ACT"))


```
checking the class of labelled variable

```{r}
class(Total_Vehicles_and_Km$Country)
levels(Total_Vehicles_and_Km$Country)
```


#Checking whether data is tidy#

Since each variable are in ts own column and each observation is in its own row and each value has a unique cell we could say the data is in tidy format.


#Subsetting $ Filtering Data#

subsetting function was used to eliminate some variables which are outside the scope of the study

```{r}
#Road_Fatalities1<-select(Road_Fatalities,-(Dayweek:Articulated.Truck..Involvement.))

```

The study focuses only on the year 2016, hence filter function usesd to filter the data for year 2016
```{r}
Road_subset <- subset(Road_Fatalities, (Year %in% c("2016")))
Road_subset

```
#Checking for missing variables/inconsistencies#

```{r}
colSums(is.na(Road_Fatalities))
```

no "NA"'s but all the unavailable information is coded as "-9", hence changing -9 to NA, then checking for missing variables.

```{r}

Road_Fatalities$Speed.Limit[Road_Fatalities$Speed.Limit == -9] <- NA
Road_Fatalities$Gender[Road_Fatalities$Gender == -9] <- NA


```

```{r}
colSums(is.na(Road_Fatalities))
```

Got one NA in Gender and the 9 "NA"'s in Speed.Limit. 

#Imputation

For eliminating NA in the Speed.Limit column, we imputed average NA's with mean value 
```{r}
Road_Fatalities$Speed.Limit[is.na(Road_Fatalities$Speed.Limit)]<-mean(Road_Fatalities$Speed.Limit,na.rm=TRUE)
sum(is.na(Road_Fatalities$Speed.Limit))
Road_Fatalities$Gender[is.na(Road_Fatalities$Gender)]<-mean(Road_Fatalities$Gender,na.rm=TRUE)
sum(is.na(Road_Fatalities$Gender))


```

NA succesfully eliminated.

#Checking for outliers#

Outliers for Speed.Limit $ Age checked using boxplot function

```{r}
boxplot((Road_Fatalities$Speed.Limit))
boxplot((Road_Fatalities$Age))

```
From boxplot, it can be predicted that there is no outliers

#Summarising $ Merging#
```{r}
new = summarise(group_by(Road_Fatalities,State),count =n())
```

For joining the data sets the variable name should be the same


```{r}
names(Total_Vehicles_and_Km)[1]<-"State"
```

```{r}
S1<-left_join(new,Total_Vehicles_and_Km,by="State")
```

#Mutate

Percentage of accidents in each state
```{r}

mutate(S1, percentageofaccidents=round((S1$count/sum(S1$count))*100,2))

```


#Data Transformation

```{r}
hist(S1$count)
hist(log10(S1$count))
```








